---
title: "graphingResults"
author: ""
date: "2023-04-28"
output: html_document
---


# SETUP

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(conflicted)
library(sna)
library(terra)
library(SpaDES)
library(data.table)
library(ggplot2)
library(readr)
library(reshape2)
library(scales)
library(tidyr)
library(RColorBrewer)
library(dplyr)
library(forcats)
library(viridis)
library(naniar)
library(raster)
library(rasterVis)
library(ggh4x)
library(egg)
library(cowplot)
library(dplyr)
library(tidyr)
library(ggspatial)
library(paletteer)
library(tidyterra)
library(gt)
library(tidyverse)
library(COINr)

options(digits = 3) # limit number of digits displayed

setPaths(
  modulePath = file.path("./modules"),
  cachePath = file.path("./cache"),
  scratchPath = file.path("./scratch"),
  inputPath = file.path("./inputs"),
  outputPath = file.path("./outputs")
)
simPaths <- getPaths()

## Get species info ####

#spList
spList <- sort(c(
  "ALFL", "BBWA", "BCCH", "BOCH", "BRCR", "CMWA", "COYE",
  "DEJU", "GCKI", "GRAJ", "LEFL", "MOWA", "OVEN", "PAWA",
  "RBNU", "RCKI", "REVI", "SWTH", "TEWA", "YRWA"
))

# facet labels sp.labs 
sp.labs <- c(
  "Alder \nFlycatcher", "Bay-breasted\n Warbler", "Black-capped\n Chickadee",
  "Boreal \nChickadee", "Brown \nCreeper", "Cape May\n Warbler",
  "Common \nYellowthroat", "Dark-eyed\n Junco", "Golden-crowned\n Kinglet",
  "Canada\n Jay", "Least \nFlycatcher", "Mourning\n Warbler",
  "Ovenbird", "Palm \nWarbler", "Red-breasted\n Nuthatch",
  "Ruby-crowned\n Kinglet", "Red-eyed\n Vireo", "Swainson’s\n Thrush",
  "Tennessee\n Warbler", "Yellow-rumped\n Warbler"
)
names(sp.labs) <- c(
  "ALFL", "BBWA", "BCCH",
  "BOCH", "BRCR", "GRAJ",
  "CMWA", "COYE", "DEJU",
  "GCKI", "LEFL", "MOWA",
  "OVEN", "PAWA", "RBNU",
  "RCKI", "REVI", "SWTH",
  "TEWA", "YRWA"
)

# spNames table
spName <- c(
  "Alder Flycatcher", "Bay-breasted Warbler", "Black-capped Chickadee",
  "Boreal Chickadee", "Brown Creeper", "Cape May Warbler",
  "Common Yellowthroat", "Dark-eyed Junco", "Golden-crowned Kinglet",
  "Canada Jay", "Least Flycatcher", "Mourning Warbler",
  "Ovenbird", "Palm Warbler", "Red-breasted Nuthatch",
  "Ruby-crowned Kinglet", "Red-eyed Vireo", "Swainson’s Thrush",
  "Tennessee Warbler", "Yellow-rumped Warbler"
)
spCode <- c(
  "ALFL", "BBWA", "BCCH",
  "BOCH", "BRCR", "CMWA",
  "COYE", "DEJU", "GCKI",
  "GRAJ", "LEFL", "MOWA",
  "OVEN", "PAWA", "RBNU",
  "RCKI", "REVI", "SWTH",
  "TEWA", "YRWA"
)
spNames <- as.data.frame(spName, spCode)

# Specify input locations ####

# studyArea
.studyAreaName <- "studyArea_AB_BCR6.shp"
locationStudyArea <- file.path(Paths$inputPath, "studyArea", tools::file_path_sans_ext(.studyAreaName), .studyAreaName)

# landscapeRasters
locationLandscapeRasters <- file.path(Paths$inputPath, "landscapeRasters")

nameForClassRas <- "vegTypesRas_AB_BCR6_2011"
locationForClass <- file.path(locationLandscapeRasters, nameForClassRas)

nameLandClassRas <- "landCoverRas_AB_BCR6_2010"
locationLandClass <- file.path(locationLandscapeRasters, nameLandClassRas)

nameAgeRas <- "ageRas_AB_BCR6_2011"
locationAge <- file.path(locationLandscapeRasters, nameAgeRas)

# rasterToMatch
rasterToMatchName <- nameLandClassRas
rasterToMatchLocation <- file.path(locationLandscapeRasters, rasterToMatchName)

# Species rasters
locationSpRas <- file.path(Paths$inputPath, "meanSpRasters")


# Specify colour schemes

#dev.new()

coloursAge <- "#A5A587FF"

agePal <- colorRampPalette(c("#D2D2C3FF", "#A5A587FF", "#78963CFF", "#1E4B3CFF", "black"))
classPalLc <- c(
  "#3C6987FF", "#87A5A5FF", "#4B3C4BFF", "#2D3C5AFF",
  "#D2D2C3FF", "#A5A587FF"
)
classPalFc <- c(
  "#B4783CFF", "#D2A54BFF",
  "#B4B45AFF", "#78963CFF", "#1E4B3CFF", "#5A695AFF"
)
classPal <- c(
  "#3C6987FF", "#87A5A5FF", "#4B3C4BFF", "#2D3C5AFF",
  "#D2D2C3FF", "#A5A587FF", "#B4783CFF", "#D2A54BFF",
  "#B4B45AFF", "#78963CFF", "#1E4B3CFF", "#5A695AFF"
)
coloursAllSame <- c(
  "#A5A587FF", "#A5A587FF", "#A5A587FF",
  "#A5A587FF", "#A5A587FF", "#A5A587FF"
)

colours2D <- classPal[10]
colours2D_alt <- classPal[12]


# coloursScale <- scale_colour_paletteer_d("impressionist.colors::la_recolte_des_foins_eragny")
# scale_color_paletteer_d("impressionist.colors::la_recolte_des_foins_eragny")
# scale_fill_paletteer_d("impressionist.colors::la_recolte_des_foins_eragny")
# classPal <- paletteer_d("impressionist.colors::la_recolte_des_foins_eragny")
#

#
# coloursLand <- classPal[1:6]
# coloursFor <- classPal[7:12]

# colours1D2D <- classPal[c(1, 10)]



# coloursAllSame2 <- c(
#   "#78963CFF", "#78963CFF", "#78963CFF",
#   "#78963CFF", "#78963CFF", "#78963CFF"
# )
```

## Specify locations of PS outputs

### Outputs for a single run 

```{r outputLocation_singleRun}
runName <- "17x10YrAgeClasses"
# runName <- "9x20yrAgeClasses"
# runName <- "5x40yrAgeClasses"
outputFolderSpPreds <- checkPath(file.path(Paths$outputPath, runName, "PSPreds"))
outputFolderSpPredsRasters <- checkPath(file.path(Paths$outputPath, runName, "PSRas"))
saveLocation <- checkPath(file.path(Paths$outputPath, "PSFigs", runName), create = TRUE)

saveLocation_1D <- checkPath(file.path(Paths$outputPath, "PSFigs"), create = TRUE)
```


### Outputs for all runs

```{r outputLocation_allRuns}
saveLocation_1D <- checkPath(file.path(Paths$outputPath, "PSFigs"), create = TRUE)

runName <- "17x10YrAgeClasses"
outputFolderSpPreds_10yr <- checkPath(file.path(Paths$outputPath, runName, "PSPreds"))
outputFolderSpPredsRasters_10yr <- checkPath(file.path(Paths$outputPath, runName, "PSRas"))
saveLocation_10yr <- checkPath(file.path(Paths$outputPath, "PSFigs", runName), create = TRUE)

runName <- "9x20yrAgeClasses"
outputFolderSpPreds_20yr <- checkPath(file.path(Paths$outputPath, runName, "PSPreds"))
outputFolderSpPredsRasters_20yr <- checkPath(file.path(Paths$outputPath, runName, "PSRas"))
saveLocation_20yr <- checkPath(file.path(Paths$outputPath, "PSFigs", runName), create = TRUE)

runName <- "5x40yrAgeClasses"
outputFolderSpPreds_40yr <- checkPath(file.path(Paths$outputPath, runName, "PSPreds"))
outputFolderSpPredsRasters_40yr <- checkPath(file.path(Paths$outputPath, runName, "PSRas"))
saveLocation_40yr <- checkPath(file.path(Paths$outputPath, "PSFigs", runName), create = TRUE)
```



## Input landscape rasters

```{r inputLandscapeRasters, echo=TRUE, message=TRUE, warning=TRUE}
# INPUT STUDY AREA and LANDSCAPE RASTERS

# get rasterToMatch
print("get rasterTomatch from local drive")
rasterToMatch <- terra::rast(rasterToMatchLocation)

# get StudyArea shapefile
print("get studyArea shapefile from local drive")
studyArea <- terra::vect(locationStudyArea)

# postProcess studyArea
studyArea <- reproducible::postProcessTo(
  from = studyArea,
  to = rasterToMatch,
  overwrite = FALSE,
  verbose = TRUE
)

# crop and mask rasterToMatch
rasterToMatch <- terra::mask(terra::crop(rasterToMatch, studyArea), studyArea)
names(rasterToMatch) <- "rasterToMatch"

# find out area
areaConversion <- 6.25
studyAreaHa <- length(na.omit(values(rasterToMatch))) * areaConversion
print(paste("The study area size is ", studyAreaHa, "ha", sep = ""))

# get forest class raster
print("get forClassRas from Drive")
forClassRas <- terra::rast(locationForClass)
forClassRas <- reproducible::postProcessTo(
      from = forClassRas,
      to = rasterToMatch,
      cropTo = studyArea,
      maskTo = studyArea,
      overwrite = FALSE,
      verbose = TRUE
    )

names(forClassRas) <- c("forClassRas")

forClassRas[forClassRas == 0] <- NA


# get non forest raster
print("get landClassRas")
landClassRas <- terra::rast(locationLandClass)
landClassRas <- reproducible::postProcessTo(
      from = landClassRas,
      to = rasterToMatch,
      cropTo = studyArea,
      maskTo = studyArea,
      overwrite = FALSE,
      verbose = TRUE
    )


# get age raster
print("get ageRas from Drive")
ageRas <- terra::rast(locationAge)
ageRas <- reproducible::postProcessTo(
      from = ageRas,
      to = rasterToMatch,
      cropTo = studyArea,
      maskTo = studyArea,
      overwrite = FALSE,
      verbose = TRUE
    )

names(ageRas) <- c("ageRas")
```


# LANDSCAPE PLOTS AND MAPS

## Summary landscape plots

```{r singlelandscapeSummaryPlots, echo=FALSE}
# Get raster values and input to cellValues
allLandData <- data.table(terra::values(c(
  forClassRas,
  landClassRas,
  ageRas,
  rasterToMatch
)))
allLandData <- setnames(allLandData, c(
  "forClass",
  "landClass",
  "age",
  "LCC05"
))

allLandData <- allLandData[, FoLRaster := ifelse(!is.na(landClass) & !is.na(forClass), "forClass",
  ifelse(!is.na(landClass), "landClass",
    ifelse(!is.na(forClass), "forClass", NA)
  )
)]

# get rid of cells with no forest or land class data.
allLandData <- allLandData[!is.na(FoLRaster)]

# make column giving the land or forest class for each row.
allLandData <- allLandData[, landForClass := coalesce(forClass, landClass)]

# in case there are classes named the same in both rasters, combine the landForClass and FoLRaster data to make sure each class is unique
allLandData <- unite(allLandData,
  landForClass,
  c(
    FoLRaster,
    landForClass
  ),
  remove = FALSE
)

allLandData$foLRaster <- as.factor(allLandData$FoLRaster)
allLandData$landForClass <- as.factor(allLandData$landForClass)
allLandData$LCC05 <- as.factor(allLandData$LCC05)

# get rid of any rows with NA values
allLandData <- subset(allLandData, !is.na(landForClass))
allLandData <- as.data.table(allLandData)

# check that there are no unwanted overlaps for nonForest raster
nonForDat <- na.omit(allLandData, cols = "landClass")
unique(nonForDat$forClass)
sort(unique(nonForDat$landForClass))
sort(unique(nonForDat$age))
sort(unique(nonForDat$FoLRaster))
sort(unique(nonForDat$LCC05))

# check that there are no unwanted overlaps for nonForest raster
forClassDat <- na.omit(allLandData, cols = "forClass")
unique(forClassDat$landClass)
sort(unique(forClassDat$landForClass))
sort(unique(forClassDat$age))
unique(forClassDat$FoLRaster)
sort(unique(forClassDat$LCC05))

landDatStats <- allLandData[
  base::order(
    FoLRaster,
    landForClass
  )
][, list(classCount = .N),
  by = list(
    FoLRaster,
    landForClass
  )
]

landDatStats[, propArea := classCount / sum(classCount)]


# Class count bar plot ####

# specify class level order
landDatStats$landForClass <- factor(landDatStats$landForClass,
  levels = c(
    "landClass_1",
    "landClass_2",
    "landClass_3",
    "landClass_4",
    "landClass_6",
    "landClass_5",
    "landClass_0",
    "forClass_3",
    "forClass_4",
    "forClass_5",
    "forClass_6",
    "forClass_1",
    "forClass_2"
  )
)


plotClassCounts <- ggplot(
  data = landDatStats,
  aes(
    x = landForClass,
    y = propArea,
  )
) +
  geom_bar(
    stat = "identity",
    position = "dodge",
    width = 0.7,
    fill = coloursAge
  ) +
  theme_minimal() +
  labs(tag = "(a)") +
  # ggtitle(paste0("Proportional area by cover class")) +
  ylab("Proportion of study area") +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_discrete(labels = c(
    "landClass_0" = "Unmanaged\n Forest",
    "landClass_1" = "Water/Ice",
    "landClass_2" = "Wetland",
    "landClass_3" = "Anthro/\nExposed",
    "landClass_4" = "Grass/\nCrop",
    "landClass_5" = "Shrub",
    "landClass_6" = "Bryoid",
    "forClass_1" = "Black \nSpruce",
    "forClass_2" = "  Black  \nSpruce Wet",
    "forClass_3" = "Deciduous",
    "forClass_4" = "Mixed",
    "forClass_5" = "Pine",
    "forClass_6" = "White \nSpruce"
  )) +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    axis.title.y = element_text(size = 12),
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    legend.text = element_blank(),
    plot.margin = margin(0, 0.5, 0, 0.1, "cm"),
    legend.position = "NULL"
  )
plotClassCounts


# Age Plot ####

forClassDat_age <- subset(forClassDat, !is.na(age))

# assign age class column
ageGrouping <- 10
maxAgeClass <- 17
maxAge <- max(forClassDat_age$age) # get age of oldest cell
maxAgeClassAge <- maxAgeClass * ageGrouping

ifelse(maxAge > maxAgeClassAge, age <- c(0:maxAge), age <- c(0:maxAgeClassAge)) # make a vector that counts from 0 to
maxAge <- as.numeric(maxAge)
noAgeClasses <- maxAge / ageGrouping
ageClass <- ifelse(noAgeClasses < maxAgeClass, ageClass <- maxAgeClass, ageClass <- noAgeClasses)
ageClass <- rep(1:ageClass, each = ageGrouping)
ageClass <- c(1, ageClass)
if (noAgeClasses > maxAgeClass) {
  ageClassDiff <- noAgeClasses - maxAgeClass
  extraClass <- maxAgeClass + (1:ageClassDiff)
  for (x in extraClass) {
    print(x)
    ageClass <- replace(ageClass, ageClass == x, maxAgeClass)
  }
}

# create a lookup table of age to age class
ageRefTab <- cbind(age, ageClass)
ageRefTab <- as.data.table(ageRefTab)
ageRefTab$age <- as.integer(ageRefTab$age)
ageRefTab$ageClass <- as.integer(ageRefTab$ageClass)
ageRefTab


forClassDat_age$age <- as.integer(forClassDat_age$age)
# storeDat <- allLandData
forClassDat_age %>% pivot_longer(cols = "age")
forClassDat_age <- forClassDat_age %>% left_join(ageRefTab, by = "age")
# allLandData %>% pivot_wider(key = ageRas, value = ageClass)
head(forClassDat_age)

forClassStats_age <- forClassDat_age[
  base::order(ageClass)
  # order the rows by the land cover class
][, list(classCount = .N),
  by = ageClass
]

forClassStats_age[, propArea := classCount / sum(classCount)]
forClassStats_age

ageDistribution <- ggplot(
  data = forClassStats_age,
  aes(
    x = ageClass,
    y = propArea
  )
) +
  geom_bar(
    stat = "identity",
    position = "dodge",
    fill = coloursAge,
    width = 0.7
  ) +
  labs(tag = "(b)") +
  # ggtitle(paste0("Forest age class structure")) +
  xlab("Age (years)") +
  ylab("Proportion of forested area") +
  theme_minimal() +
  scale_x_discrete(
    labels = c(
      "1" = "0-10", "2" = "", "3" = "21-30", "4" = "",
      "5" = "41-50", "6" = "", "7" = "61-70", "8" = "",
      "9" = "81-90", "10" = "", "11" = "101-110", "12" = "",
      "13" = "121-130", "14" = "", "15" = "141-150", "16" = "",
      "17" = "161+"
    ),
    limits = c(
      "1", "2", "3", "4", "5", "6", "7", "8", "9", "10",
      "11", "12", "13", "14", "15", "16", "17"
    )
  ) +
  scale_y_continuous(breaks = pretty_breaks()) +
  theme(
    title = element_text(size = 14),
    strip.text.x = element_text(size = 12),
    strip.text.y = element_text(size = 12),
    strip.background = element_rect(
      color = "white", fill = "white", linewidth = 1.5
    ),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    axis.title.y = element_text(size = 12),
    axis.title.x = element_text(size = 12),
    legend.title = element_blank(),
    legend.text = element_blank(),
    plot.margin = margin(0, 0, 0.5, 0.3, "cm"),
    legend.position = "NULL"
  )

ageDistribution


# Bar plot of mean age per landForClass ####


ageStats <- forClassDat[
  base::order(landForClass)
][, list(
  classCount = .N,
  meanAge = mean(age)
),
by = landForClass
]

# specify class level order
ageStats$landForClass <- factor(ageStats$landForClass,
  levels = c(
    "landClass_1",
    "landClass_2",
    "landClass_3",
    "landClass_4",
    "landClass_6",
    "landClass_5",
    "landClass_0",
    "forClass_3",
    "forClass_4",
    "forClass_5",
    "forClass_6",
    "forClass_1",
    "forClass_2"
  )
)


plotMeanAgeByClass <- ggplot(
  data = ageStats,
  aes(
    x = landForClass,
    y = meanAge
  )
) +
  geom_bar(
    stat = "identity",
    position = "dodge",
    width = 0.7,
    fill = coloursAge
  ) +
  theme_minimal() +
  # ggtitle(paste0("Mean age by forest class")) +
  labs(tag = "(c)") +
  xlab("Forest class") +
  ylab("Mean Age (years)") +
  scale_x_discrete(labels = c(
    "forClass_1" = "Black \nSpruce",
    "forClass_2" = "  Black  \nSpruce Wet",
    "forClass_3" = "Deciduous",
    "forClass_4" = "Mixed",
    "forClass_5" = "Pine",
    "forClass_6" = "White \nSpruce"
  )) +
  theme(
    title = element_text(size = 14),
    strip.text.x = element_text(size = 12),
    strip.text.y = element_text(size = 12),
    strip.background = element_rect(
      color = "white", fill = "white", linewidth = 1.5
    ),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    axis.title.y = element_text(size = 12),
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 10),
    plot.margin = margin(0, 0.5, 0, 0, "cm"),
    legend.position = "none"
  )
plotMeanAgeByClass


# Make single landscape plot figure ####

lay <- rbind(c(1, 1), c(2, 3))
landscapeSummary <- gridExtra::grid.arrange(plotClassCounts,
  ageDistribution,
  plotMeanAgeByClass,
  layout_matrix = lay
)

ggsave(file.path(saveLocation_1D, "landscapeSummary.jpeg"), landscapeSummary, width = 16, height = 20, units = "cm")
```


## landscape rasters maps


```{r singleStudyAreaLandscapeRasterMaps, echo=FALSE}

op <- par(cex = 2) # change legend text size

# forClassRas

# reorder classes
is <- c(1, 2, 3, 4, 5, 6)
becomes <- c(11, 12, 7, 8, 9, 10)
reclassMatrixFcr <- matrix(c(is, becomes),
  ncol = 2,
  byrow = FALSE
)
print(reclassMatrixFcr)
forClassRas <- terra::classify(forClassRas, reclassMatrixFcr)

# reclassify to Class Names
fcrValue <- c(7, 8, 9, 10, 11, 12)
fcrLabel <- c("7", "8", "9", "10", "11", "12")
fcrLabelReclass <- c("Deciduous", "Mixed", "Pine", "White Spruce", "Black Spruce", "Black Spruce Wet")
fcrReclassified <- data.frame(fcrValue, fcrLabel, fcrLabelReclass)
fcrReclassified$fcrLabelReclass <- as.factor(fcrReclassified$fcrLabelReclass)
print(fcrReclassified)
str(fcrReclassified)

forClassRas <- categories(forClassRas, value = fcrReclassified, active = 2)

studyArea <- terra::project(studyArea, y = "+proj=longlat +datum=WGS84")
forClassRas <- terra::project(forClassRas, y = "+proj=longlat +datum=WGS84")
forClassRas <- crop(forClassRas, studyArea)

dev.new()
plot(forClassRas, legend = TRUE, col = classPalFc, colNA = "white", axes = FALSE, box = FALSE)



# landClassRas

# reorder classes
is <- c(0, 1, 2, 3, 4, 5, 6)
becomes <- c(13, 7, 8, 9, 10, 11, 12)
reclassMatrixLcr <- matrix(c(is, becomes),
  ncol = 2,
  byrow = FALSE
)
print(reclassMatrixLcr)
landClassRas <- terra::classify(landClassRas, reclassMatrixLcr)

# reclassify to Class Names
lcrValue <- c(13, 7, 8, 9, 10, 11, 12)
lcrLabel <- c("13", "7", "8", "9", "10", "11", "12")
lcrLabelReclass <- c("Unmanaged Forest", "Water/Ice", "Wetland", "Anthro/Exposed Land", "Grass/Cropland", "Shrub", "Bryoid")
lcrReclassified <- data.frame(lcrValue, lcrLabel, lcrLabelReclass)
lcrReclassified$lcrLabelReclass <- as.factor(lcrReclassified$lcrLabelReclass)
print(lcrReclassified)
str(lcrReclassified)

landClassRas <- categories(landClassRas, value = lcrReclassified, active = 2)

landClassRas <- terra::project(landClassRas, y = "+proj=longlat +datum=WGS84")
landClassRas <- crop(landClassRas, studyArea)

names(landClassRas) <- c("landClassRas")
landClassRas <- terra::mask(landClassRas, forClassRas, inverse = TRUE)

dev.new()
plot(landClassRas, legend = TRUE, col = classPalLc, colNA = "white", axes = FALSE, box = FALSE)



# ageRas

ageRas <- terra::project(ageRas, y = "+proj=longlat +datum=WGS84")
ageRas <- crop(ageRas, studyArea)

dev.new()
plot(ageRas, colNA = "white", axes = FALSE, box = FALSE, col = agePal(75))


```


#MAKE TABLES WITH SP AND CLASS INFO

### Sp Info Table
```{r spInfo, echo=FALSE}
commonName <- c("Alder Flycatcher", "Bay-breasted Warbler", "Black-capped Chickadee", "Boreal Chickadee", "Brown Creeper", "Canada Jay", "Cape May Warbler", "Common Yellowthroat", "Dark-eyed Junco", "Golden-crowned Kinglet", "Least Flycatcher", "Mourning Warbler", "Ovenbird", "Palm Warbler", "Red-breasted Nuthatch", "Ruby-crowned Kinglet", "Red-eyed Vireo", "Swainson’s Thrush", "Tennessee Warbler", "Yellow-rumped Warbler")

scientificName <- c(
  "Empidonax alnorum", "Setophaga castanea", "Poecile atricapillus", "Poecile hudsonicus", "Certhia americana", "Perisoreus canadensis", "Setophaga tigrina", "Geothlypis trichas",
  "Junco hyemalis", "Regulus satrapa", "Empidonax minimus", "Geothlypis philadelphia", "Seiurus aurocapilla", "Setophaga palmarum", "Sitta canadensis", "Regulus calendula", "Vireo olivaceus", "Catharus ustulatus", "Leiothlypis peregrina", "Setophaga coronata"
)

spInfo <- data.frame(commonName, scientificName)
names(spInfo) <- c("Common Name", "Scientific Name")

spInfoTab <- gt::gt(spInfo) |>
  tab_style(
    locations = list(
      cells_body(
        columns = "Scientific Name", rows = c(1:20)
      )
    ),
    style = list(
      cell_text(style = "italic")
    )
  )

spInfoTab

gtsave(spInfoTab, "spInfoTab.png", path = saveLocation_1D)
```



### Class Info Table
```{r classInfo, echo=FALSE}
Class <- c(
  "Water/Ice",
  "Wetland",
  "Anthropogenic/Exposed Land",
  "Grass/Cropland",
  "Shrubland",
  "Bryoid",
  "Unmanaged Forest",
  "Deciduous",
  "Mixedwood",
  "Pine",
  "White Spruce",
  "Black Spruce",
  "Black Spruce Wetland"
)


Description <- c(
  "Lakes and rivers, or consistent snow/ice cover throughout the year.",
  "Areas of poor drainage, with vegetation but less than 5% tree cover.",
  "Non-forested land, no cover of any type, including rock and urban environments.",
  "Non-forested land. Herbaceous grassland or herbaceous cover.",
  "Non-forested land. Open shrubs or closed shrubs.",
  "Non-forested land. Bryoid cover.",
  "Miscellaneous forest areas that are considered forest on the land cover class raster, but not the forest class raster. This class is additional to those described in Lane-Shaw, Raymundo and Cumming (2020).",
  "Stands where combined trembling aspen, balsam poplar and white birch comprise more than 80%.",
  "Stands where deciduous species comprise more than 20% and combined conifer species (jack pine, lodgepole pine, white spruce, balsam fir, and black spruce) more than 20%.",
  "Stands where combined jack pine and lodgepole pine are the leading species and deciduous species comprise less than 20%",
  "Stands where combined white spruce and balsam fir comprise more than 80%.",
  "Stands where black spruce is the leading species and larch = 0%, or stands where black spruce is the leading species and the combined species trembling aspen + balsam poplar + balsam fir + jack pine comprise more than 0%.",
  "Stands where black spruce is the leading species, or stands where black spruce is the leading species and larch comprises more than 0%, or stands where black spruce is the leading species and the combined species trembling aspen + balsam poplar + balsam fir + jack pine = 0."
)

classInfo <- data.frame(Class, Description)


classInfoTab <- gt::gt(classInfo) |>
  tab_row_group(
    label = "Forest Classes",
    rows = c(8:13)
  ) |>
  tab_row_group(
    label = "Land Cover Classes",
    rows = c(1:7)
  )

classInfoTab
gtsave(classInfoTab, "classInfoTab.png", path = saveLocation_1D, expand = 2000)
```


# PLOT MAIN 1D PREDICTION RESULTS

###make plot of 1D mean density by cover class 

```{r meanDensity1D, echo=FALSE}

# make plot of 1D mean density by cover class
spPreds1D <- lapply(spList, FUN = function(sp) {
  print(sp)
  spPreds1D <- as.data.frame(read_csv(paste(outputFolderSpPreds, "/", sp, "_spPreds1D.csv", sep = "")))

  spPreds1D$landForClass <- as.factor(spPreds1D$landForClass)
  spPreds1D$FoLRaster <- as.factor(spPreds1D$FoLRaster)
  spPreds1D$sp <- as.factor(spPreds1D$sp)
  return(spPreds1D)
})
names(spPreds1D) <- spList

spPreds1DSingleFrame <- rbindlist(spPreds1D)

spPreds1DSingleFrame <- rbindlist(spPreds1D[1:20])


examine1DPreds <- spPreds1DSingleFrame[, list(
  meanDensity = mean(meanSpDensity),
  medianDensity = median(meanSpDensity),
  maxDensity = max(meanSpDensity),
  minDensity = min(meanSpDensity)
),
by = sp
][
  base::order(maxDensity)
]

# specify class level order
spPreds1DSingleFrame$landForClass <- factor(spPreds1DSingleFrame$landForClass,
  levels = c(
    "landClass_1",
    "landClass_2",
    "landClass_3",
    "landClass_4",
    "landClass_6",
    "landClass_5",
    "landClass_0",
    "forClass_3",
    "forClass_4",
    "forClass_5",
    "forClass_6",
    "forClass_1",
    "forClass_2"
  )
)


# Make plot
plotMeanSpDensity <- ggplot(
  data = spPreds1DSingleFrame,
  aes(
    x = landForClass,
    y = meanSpDensity,
    fill = landForClass
  )
) +
  facet_wrap(~sp, 
    labeller = labeller(sp = sp.labs),
    ncol = 4
  ) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(
    values = classPal,
    labels = c(
      "landClass_0" = "Unmanaged \nForest",
      "landClass_1" = "Water/Ice",
      "landClass_2" = "Wetland",
      "landClass_3" = "Anthro/\nExposed",
      "landClass_4" = "Grass/\nCrop",
      "landClass_5" = "Shrub",
      "landClass_6" = "Bryoid",
      "forClass_1" = "Black \nSpruce",
      "forClass_2" = "Black \nSpruce Wet",
      "forClass_3" = "Deciduous",
      "forClass_4" = "Mixed",
      "forClass_5" = "Pine",
      "forClass_6" = "White \nSpruce"
    )
  ) +
  geom_errorbar(
    aes(
      ymin = meanSpDensity - seSpDensity,
      ymax = meanSpDensity + seSpDensity
    ),
    width = .5
  ) +
  theme_minimal() +
  ggtitle("Mean density predictions by forest or land cover class") +
  xlab("Land or Forest Class") +
  ylab("Mean Predicted Density (males per ha)") +
  scale_y_continuous(breaks = pretty_breaks()) +
  theme(
    title = element_blank(),
    strip.text.x = element_text(
      size = 15
    ),
    strip.text.y = element_text(
      size = 15
    ),
    legend.key.size = unit(0.3, units = "cm"),
    legend.title = element_blank(),
    legend.byrow = TRUE,
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    strip.background = element_rect(
      color = "white", fill = "white", linewidth = 0.001
    ),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 18),
    axis.title.x = element_text(size = 18),
    plot.margin = margin(0.3, 0.2, 0.2, 0.2, "cm")
  )


plotMeanSpDensity

# save
ggsave(file.path(saveLocation_1D, "1DResults.jpeg"), plotMeanSpDensity, width = 20, height = 37, units = "cm")

```


# PLOT 2D PREDICTION RESULTS


####### 10 yr age classes

```{r plotSpMatricies, echo=FALSE}
# make plot of 2D sp density predictions

# get matricies
spMatricies <- lapply(spList, FUN = function(sp) {
 
    spMatrix <- as.data.frame(read_csv(paste(outputFolderSpPreds_10yr, "/", sp, "_matrix.csv", sep = "")))
  print(spMatrix)
  return(spMatrix)
})
names(spMatricies) <- spList


matrixAsTables <- lapply(X = spList, FUN = function(sp) {

  matrix <- eval(parse(text = paste("spMatricies$", sp, sep = "")))
  matrix <- matrix[2:ncol(matrix)]
  matrix <- data.matrix(matrix)
  tab <- melt(matrix)
  colnames(tab) <- c("forClass", "ageClass", "spDensityPred")
  species <- rep(paste(sp), nrow(tab))
  maxDensity <- rep(max(tab$spDensityPred), nrow(tab))
  tab <- cbind(tab, species = species, maxDensity = maxDensity)
  tab$ageClass <- as.integer(tab$ageClass)
  tab$forClass <- as.character(tab$forClass)
  tab$species <- as.character(tab$species)
  tab$spDensityPred <- as.numeric(tab$spDensityPred)

  print(tab)

  return(tab)
})

names(matrixAsTables) <- spList

dataFrameForGraph <- rbindlist(matrixAsTables)
dataFrameForGraph <- dataFrameForGraph[base::order(maxDensity)]
speciesInOrder <- unique(dataFrameForGraph$species)

# make dataframe of sps you want to include
spsWanted <- c(speciesInOrder[1:10])
# spsWanted <- speciesInOrder[11:20]
dataFrameForGraph <- dataFrameForGraph[dataFrameForGraph[["species"]] %in% spsWanted]

# specify class level order
dataFrameForGraph$species <- factor(dataFrameForGraph$species,
  levels = spsWanted
)
dataFrameForGraph$forClass <- factor(dataFrameForGraph$forClass,
  levels = c(
    "3",
    "4",
    "5",
    "6",
    "1",
    "2"
  )
)

# facet label names for forest classes
class.labs <- c("Black \nSpruce", "Black \nSpruce Wet", "Deciduous", "Mixed", "Pine", "White \nSpruce")
names(class.labs) <- c("1", "2", "3", "4", "5", "6")

#plot
plotMeanSpDensity2D <- ggplot(
  data = dataFrameForGraph,
  aes(y = spDensityPred, x = ageClass, fill = forClass)
) + 
  geom_bar(position = "dodge", stat = "identity") +
  facet_grid(species ~ forClass,
    labeller = labeller(forClass = class.labs, species = sp.labs)
  ) +
  theme_minimal() +
  scale_fill_manual(values = coloursAllSame) +
  xlab("Forest Age (years)") +
  scale_x_discrete(
    labels = c(
      "1" = "0-10",
      "5" = "41-50",
      "9" = "81-90",
      "13" = "121-130",
      "17" = "161+"
    ),
    limits = c(
      "1", "2", "3", "4",
      "5", "6", "7", "8",
      "9", "10", "11", "12",
      "13", "14", "15", "16", "17"
    ),
    breaks = c("1", "5", "9", "13", "17")
  ) +
  ylab("Predicted Density (males per ha)") +
  scale_y_continuous(breaks = pretty_breaks(n = 4)) + 
  theme(
    title = element_blank(),
    strip.text.x = element_text(
      size = 10
    ),
    strip.text.y = element_text(
      size = 10
    ),
    legend.position = "none",
    strip.background = element_rect(
      color = "white", fill = "white", linewidth = 1.1
    ),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    axis.title.y = element_text(size = 12),
    axis.title.x = element_text(size = 12),
    plot.margin = margin(0.2, 0.1, 0.1, 0.1, "cm")
  )


ggsave(file.path(saveLocation_10yr, "2DResults1.jpeg"), plotMeanSpDensity2D, width = 22, height = 30, units = "cm")


```


###### 20 yr age classes

```{r plotSpMatricies_20, echo=FALSE}
# make plot of 2D sp density predictions

# get matricies
spMatricies <- lapply(spList, FUN = function(sp) {
   spMatrix <- as.data.frame(read_csv(paste(outputFolderSpPreds_20yr, "/", sp, "_matrix.csv", sep = "")))
   return(spMatrix)
})
names(spMatricies) <- spList


matrixAsTables <- lapply(X = spList, FUN = function(sp) {
  matrix <- eval(parse(text = paste("spMatricies$", sp, sep = "")))
  matrix <- matrix[2:ncol(matrix)]
  matrix <- data.matrix(matrix)
  tab <- melt(matrix)
  colnames(tab) <- c("forClass", "ageClass", "spDensityPred")
  species <- rep(paste(sp), nrow(tab))
  maxDensity <- rep(max(tab$spDensityPred), nrow(tab))
  tab <- cbind(tab, species = species, maxDensity = maxDensity)
  tab$ageClass <- as.integer(tab$ageClass)
  tab$forClass <- as.character(tab$forClass)
  tab$species <- as.character(tab$species)
  tab$spDensityPred <- as.numeric(tab$spDensityPred)

  print(tab)

  return(tab)
})

names(matrixAsTables) <- spList

dataFrameForGraph <- rbindlist(matrixAsTables)
dataFrameForGraph <- dataFrameForGraph[base::order(maxDensity)]
speciesInOrder <- unique(dataFrameForGraph$species)

# make dataframe of sps you want to include
spsWanted <- speciesInOrder[1:10]
# spsWanted <- speciesInOrder[11:20]
dataFrameForGraph <- dataFrameForGraph[dataFrameForGraph[["species"]] %in% spsWanted]

# specify class level order
dataFrameForGraph$species <- factor(dataFrameForGraph$species,
  levels = spsWanted
)
dataFrameForGraph$forClass <- factor(dataFrameForGraph$forClass,
  levels = c(
    "3",
    "4",
    "5",
    "6",
    "1",
    "2"
  )
)

# facet label names for forest classes
class.labs <- c("Black \nSpruce", "Black \nSpruce Wet", "Deciduous", "Mixed", "Pine", "White \nSpruce")
names(class.labs) <- c("1", "2", "3", "4", "5", "6")

# plot
plotMeanSpDensity2D <- ggplot(
  data = dataFrameForGraph,
  aes(y = spDensityPred, x = ageClass, fill = forClass)
) + #
  geom_bar(position = "dodge", stat = "identity") +
  facet_grid(species ~ forClass,
    labeller = labeller(forClass = class.labs, species = sp.labs)
  ) +
  theme_minimal() +
  scale_fill_manual(values = coloursAllSame) +
  xlab("Forest Age (years)") +
  scale_x_discrete(
    labels = c(
      "1" = "0-20",
      "3" = "41-60",
      "5" = "81-100",
      "7" = "121-140",
      "9" = "161+"
    ),
    limits = c(
      "1", "2", "3", "4",
      "5", "6", "7", "8",
      "9"
    ),
    breaks = c("1", "3", "5", "7", "9")
  ) +
  ylab("Predicted density (male sps per hectare)") +
  scale_y_continuous(breaks = pretty_breaks(n = 4)) + 
  theme(
    title = element_blank(),
    strip.text.x = element_text(
      size = 10
    ),
    strip.text.y = element_text(
      size = 10
    ),
    legend.position = "none",
    strip.background = element_rect(
      color = "white", fill = "white", linewidth = 1.1
    ),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    axis.title.y = element_text(size = 12),
    axis.title.x = element_text(size = 12),
    plot.margin = margin(0.2, 0.1, 0.1, 0.1, "cm")
  )

ggsave(file.path(saveLocation_20yr, "2DResults1.jpeg"), plotMeanSpDensity2D, width = 22, height = 30, units = "cm")

plotMeanSpDensity2D
```

###### 40 yr age classes

```{r plotSpMatricies_40, echo=FALSE}
# make plot of 2D sp density predictions

# get matricies
spMatricies <- lapply(spList, FUN = function(sp) {
  spMatrix <-  as.data.frame(read_csv(paste(outputFolderSpPreds_40yr, "/", sp, "_matrix.csv", sep = "")))
   return(spMatrix)
})
names(spMatricies) <- spList


matrixAsTables <- lapply(X = spList, FUN = function(sp) {
    matrix <- eval(parse(text = paste("spMatricies$", sp, sep = "")))
  matrix <- matrix[2:ncol(matrix)]
  matrix <- data.matrix(matrix)
  tab <- melt(matrix)
  colnames(tab) <- c("forClass", "ageClass", "spDensityPred")
  species <- rep(paste(sp), nrow(tab))
  maxDensity <- rep(max(tab$spDensityPred), nrow(tab))
  tab <- cbind(tab, species = species, maxDensity = maxDensity)
  tab$ageClass <- as.integer(tab$ageClass)
  tab$forClass <- as.character(tab$forClass)
  tab$species <- as.character(tab$species)
  tab$spDensityPred <- as.numeric(tab$spDensityPred)

  print(tab)

  return(tab)
})

names(matrixAsTables) <- spList

dataFrameForGraph <- rbindlist(matrixAsTables)
dataFrameForGraph <- dataFrameForGraph[base::order(maxDensity)]
speciesInOrder <- unique(dataFrameForGraph$species)

# make dataframe of sps you want to include
spsWanted <- c(speciesInOrder[1:10])
#spsWanted <- speciesInOrder[11:20]

dataFrameForGraph <- dataFrameForGraph[dataFrameForGraph[["species"]] %in% spsWanted]

# specify class level order
dataFrameForGraph$species <- factor(dataFrameForGraph$species,
  levels = spsWanted
)
dataFrameForGraph$forClass <- factor(dataFrameForGraph$forClass,
  levels = c(
    "3",
    "4",
    "5",
    "6",
    "1",
    "2"
  )
)

# facet label names for forest classes
class.labs <- c("Black \nSpruce", "Black \nSpruce Wet", "Deciduous", "Mixed", "Pine", "White \nSpruce")
names(class.labs) <- c("1", "2", "3", "4", "5", "6")

# plot
plotMeanSpDensity2D <- ggplot(
  data = dataFrameForGraph,
  aes(y = spDensityPred, x = ageClass, fill = forClass)
) + 
  geom_bar(position = "dodge", stat = "identity") +
  facet_grid(species ~ forClass,
    labeller = labeller(forClass = class.labs, species = sp.labs)
  ) +
  theme_minimal() +
  scale_fill_manual(values = coloursAllSame) +
  xlab("Forest Age (years)") +
  scale_x_discrete(
    labels = c(
      "1" = "0-40",
      "2" = "41-80",
      "3" = "81-120",
      "4" = "121-160",
      "5" = "161+"
    ),
    limits = c("1", "2", "3", "4", "5"),
    breaks = c("1", "2", "3", "4", "5")
  ) +
  ylab("Predicted density (male birds per hectare)") +
  scale_y_continuous(breaks = pretty_breaks(n = 4)) + theme(
    title = element_blank(),
    strip.text.x = element_text(
      size = 10
    ),
    strip.text.y = element_text(
      size = 10
    ),
    legend.position = "none",
    strip.background = element_rect(
      color = "white", fill = "white", linewidth = 1.1
    ),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    axis.title.y = element_text(size = 12),
    axis.title.x = element_text(size = 12),
    plot.margin = margin(0.2, 0.1, 0.1, 0.1, "cm")
  )

ggsave(file.path(saveLocation_40yr, "/2DResults1.jpeg"), plotMeanSpDensity2D, width = 22, height = 30, units = "cm")

plotMeanSpDensity2D
```


# TABLES OF 2D PREDICTIONS

### 10-year age classes

```{r plotSpMatriciesTables, echo=FALSE}
# make 2D sp density predictions table

# get matricies
gtTabs <- lapply(spList, FUN = function(sp) {
  print(sp)
  nameSp <- spNames[sp, ]

  spMatrix <- as.data.frame(read_csv(paste(outputFolderSpPreds_10yr, "/", sp, "_matrix.csv", sep = "")))
    names(spMatrix) <- c("forestClass",
                       "0-10", "11-20", "21-30", "31-40", "41-50",
                       "51-60", "61-70", "71-80", "81-90", "91-100", 
                       "101-110", "111-120", "121-130", "131-140", 
                       "141-150", "151-160", "161+")

  spMatrix$forestClass[spMatrix$forestClass == 1] <- "Black Spruce"
  spMatrix$forestClass[spMatrix$forestClass == 2] <- "Black Spruce Wet"
  spMatrix$forestClass[spMatrix$forestClass == 3] <- "Deciduous"
  spMatrix$forestClass[spMatrix$forestClass == 4] <- "Mixed"
  spMatrix$forestClass[spMatrix$forestClass == 5] <- "Pine"
  spMatrix$forestClass[spMatrix$forestClass == 6] <- "White Spruce"


  spMatrix$forestClass <- as.factor(spMatrix$forestClass)
  spMatrix$forestClass <- factor(spMatrix$forestClass, levels = c("Deciduous", "Mixed", "Pine", "White Spruce", "Black Spruce", "Black Spruce Wet"))
  # Define the new order for the Class column
  new_order <- c("Deciduous", "Mixed", "Pine", "White Spruce", "Black Spruce", "Black Spruce Wet")
  # Reorder the rows based on the new order
  spMatrix <- spMatrix[match(new_order, spMatrix$forestClass), ]


  spMatrix <- COINr::signif_df(spMatrix, digits = 2)
  names(spMatrix) <- c("forestClass", "0-10", "11-20", "21-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90", "91-100", "101-110", "111-120", "121-130", "131-140", "141-150", "151-160", "161+")


  gtTab <- gt::gt(spMatrix, rowname_col = "forestClass") |> 
    tab_header(title = paste("Two-factor Piecewise Smoothing Predictions (males per ha) for ", nameSp, sep = "")) |>
    tab_stubhead(label = "Forest Class") |>
    tab_spanner(
      label = "Forest Age Class (Years)",
      columns = c("0-10", "11-20", "21-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90", "91-100", "101-110", "111-120", "121-130", "131-140", "141-150", "151-160", "161+")
    )
  
  return <- gtTab
})

names(gtTabs) <- spList

gtTabs$OVEN

```


### All piecewise smoothing results in a single table (all age classes)

```{r plotSpMatriciesTables_all, echo=FALSE}
# make 2D sp density predictions table

# get matricies
lapply(spList, FUN = function(sp) {
 
  print(sp)
  nameSp <- spNames[sp, ]
 
  # prep 1DPS preds
  spPreds1D <- as.data.frame(read_csv(paste(outputFolderSpPreds_10yr, "/", sp, "_spPreds1D.csv", sep = "")))
  spPreds1D <- spPreds1D[spPreds1D$FoLRaster == "forClass", ]
  spPreds1D <- spPreds1D[, c("landForClass", "meanSpDensity")]
  spPreds1D <- pivot_wider(spPreds1D, names_from = "landForClass", values_from = "meanSpDensity")
  ageClass <- "All Ages"
  Method <- "1DPS"
  spPreds1D <- cbind(Method, ageClass, spPreds1D)
  names(spPreds1D) <- c("Method", "ageClass", "1", "2", "3", "4", "5", "6")
  print(spPreds1D)

  # prep 2DPS preds (10 yr)
  spMatrix_10 <- as.data.frame(read_csv(paste(outputFolderSpPreds_10yr, "/", sp, "_matrix.csv", sep = "")))
  print(spMatrix_10)
  names(spMatrix_10) <- c("forestClass", "0-10", "11-20", "21-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90", "91-100", "101-110", "111-120", "121-130", "131-140", "141-150", "151-160", "161+")
  spMatrix_10$forestClass <- as.factor(spMatrix_10$forestClass)
  spMatrix_10 <- pivot_longer(spMatrix_10,
    cols = c(2:18),
    names_to = "ageClass", 
    values_to = "value"
  )
  spMatrix_10 <- pivot_wider(spMatrix_10, names_from = "forestClass", values_from = "value")
  Method <- rep("2DPS (10-year age classes)", times = nrow(spMatrix_10))
  spMatrix_10 <- cbind(Method, spMatrix_10)
  print(spMatrix_10)

   # prep 2DPS preds (20 yr)
  spMatrix_20 <- as.data.frame(read_csv(paste(outputFolderSpPreds_20yr, "/", sp, "_matrix.csv", sep = "")))
  names(spMatrix_20) <- c("forestClass", "0-20", "21-40", "41-60", "61-80", "81-100", "101-120", "121-140", "141-160", "161+")
  spMatrix_20$forestClass <- as.factor(spMatrix_20$forestClass)
  spMatrix_20 <- pivot_longer(spMatrix_20,
    cols = c(2:10),
    names_to = "ageClass", 
    values_to = "value"
  )
  spMatrix_20 <- pivot_wider(spMatrix_20, names_from = "forestClass", values_from = "value")
  Method <- rep("2DPS (20-year age classes)", times = nrow(spMatrix_20))
  spMatrix_20 <- cbind(Method, spMatrix_20)
  print(spMatrix_20)

   # prep 2DPS preds (40 yr)
  spMatrix_40 <- as.data.frame(read_csv(paste(outputFolderSpPreds_40yr, "/", sp, "_matrix", ".csv", sep = "")))
    names(spMatrix_40) <- c("forestClass", "0-40", "41-80", "81-120", "121-160", "161+")
  spMatrix_40$forestClass <- as.factor(spMatrix_40$forestClass)
  spMatrix_40 <- pivot_longer(spMatrix_40,
    cols = c(2:6),
    names_to = "ageClass", # new column for day names
    values_to = "value"
  )
  spMatrix_40 <- pivot_wider(spMatrix_40, names_from = "forestClass", values_from = "value")
  Method <- rep("2DPS (40-year age classes)", times = nrow(spMatrix_40))
  spMatrix_40 <- cbind(Method, spMatrix_40)
print(spMatrix_40)
  
#put all preds together
  spMatrix <- rbind(spPreds1D, spMatrix_40, spMatrix_20, spMatrix_10)
  names(spMatrix) <- c("Method", "Age Class", "Black Spruce", "Black Spruce Wet", "Deciduous", "Mixed", "Pine", "White Spruce")
  spMatrix <- spMatrix[, c("Method", "Age Class", "Deciduous", "Mixed", "Pine", "White Spruce", "Black Spruce", "Black Spruce Wet")]

  spMatrix <- COINr::signif_df(spMatrix, digits = 2)
  names(spMatrix) <- c("Method", "Age Class", "Deciduous", "Mixed", "Pine", "White Spruce", "Black Spruce", "Black Spruce Wet")
print(spMatrix)
   
  #save csv of all results
  fileName <- paste(sp, "_allPSPredictions", sep = "")
  write.csv(spMatrix, file = file.path(saveLocation_1D, fileName))
 
  #make table
  gtTab <- gt::gt(spMatrix, rowname_col = "Age Class") |> 
    tab_header(title = paste("Piecewise Smoothing Predictions for ", nameSp, sep = "")) |>
    tab_stubhead(label = " Forest Age (Years)") |>
    tab_spanner(
      label = "Forest Class",
      columns = c("Deciduous", "Mixed", "Pine", "White Spruce", "Black Spruce", "Black Spruce Wet")
    ) |>
    tab_row_group(
      label = "1D Piecewise Smoothing",
      rows = 1
    ) |>
    tab_row_group(
      label = "2DPS with 40-year age classes",
      rows = c(2:6)
    ) |>
    tab_row_group(
      label = "2DPS with 20-year age classes",
      rows = c(7:15)
    ) |>
    tab_row_group(
      label = "2DPS with 10-year age classes",
      rows = c(16:32)
    )
  
  gtsave(gtTab, paste(sp, "_AllPSResults.png", sep = ""), path = saveLocation_1D, expand = 50)
})


```

# DENSITY PLOTS

### Histograms of 1D class data 

```{r histogram1D, echo=FALSE}
class.labs <- c("Unmanaged Forest", "Water/Ice", "Wetland", "Anthro/Exposed", "Grass/Crop", "Shrub", "Bryoid", "Black Spruce", "Black Spruce Wet", "Deciduous", "Mixed", "Pine", "White Spruce")
names(class.labs) <- c("landClass_0", "landClass_1", "landClass_2", "landClass_3", "landClass_4", "landClass_5", "landClass_6", "forClass_1", "forClass_2", "forClass_3", "forClass_4", "forClass_5", "forClass_6")


lapply(spList, FUN = function(sp) {
  
  print(sp)
  nameSp <- spNames[sp, ]
 
  spDatasets <- as.data.frame(read_csv(paste(outputFolderSpPreds, "/", sp, "_fullDataset.csv", sep = "")))
    spDatasets <- spDatasets %>% tidyterra::select(-1)

  names(spDatasets) <- c(
    "spDensity", "forClass",
    "landClass", "age", "landForClass", "FoLRaster", "sp", "studyArea"
  )
  spDatasets$landForClass <- as.factor(spDatasets$landForClass)
  spDatasets$sp <- as.factor(spDatasets$sp)
  spDatasets$age <- as.integer(spDatasets$age)
  spDatasets$landForClass <- factor(spDatasets$landForClass,
    levels = c(
      "landClass_1",
      "landClass_2",
      "landClass_3",
      "landClass_4",
      "landClass_6",
      "landClass_5",
      "landClass_0",
      "forClass_3",
      "forClass_4",
      "forClass_5",
      "forClass_6",
      "forClass_1",
      "forClass_2"
    )
  )


  spHistogram <- ggplot(spDatasets, aes(x = spDensity)) + 
    geom_histogram(fill = coloursAge) +
    facet_wrap(~landForClass,
      labeller = labeller(landForClass = class.labs),
      ncol = 4
    ) +
     theme_minimal() +
    #    ggtitle(paste("Histograms of ", nameSp, " density data \nby forest or land cover class", sep = "")) +
    xlab("BAM predicted density (males ha⁻¹)") +
    scale_x_continuous(breaks = pretty_breaks(n = 5), limits = c(0, NA)) +
    ylab("Count") +
    scale_y_continuous(breaks = pretty_breaks(n = 5)) +
    theme(
      # title = element_text(size = 14),
      title = element_blank(),
      strip.text.x = element_text(
        size = 14),
      strip.background = element_rect(
        color = "white", fill = "white", linewidth = 1.1),
      legend.title = element_blank(),
      legend.position = "none",
      axis.text.y = element_text(size = 12),
      axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
      axis.title = element_text(size = 14)
    )
  
  print(spHistogram)
  
  #save
  ggsave(file.path(saveLocation_1D, paste(sp, "_histogram.jpeg", sep = "")), spHistogram, width = 20, height = 20, units = "cm")
})
```


### Kernel density plot of 1D data 

Creates a density plot for a single species, selected from the spDatasets, for each class. 

```{r densityPlot1D, echo=FALSE}

class.labs <- c("Unmanaged Forest", "Water/Ice", "Wetland", "Anthro/Exposed", "Grass/Crop", "Shrub", "Bryoid", "Black Spruce", "Black Spruce Wet", "Deciduous", "Mixed", "Pine", "White Spruce")
names(class.labs) <- c("landClass_0", "landClass_1", "landClass_2", "landClass_3", "landClass_4", "landClass_5", "landClass_6", "forClass_1", "forClass_2", "forClass_3", "forClass_4", "forClass_5", "forClass_6")

# get spDatasets

lapply(spList, FUN = function(sp) {
 
  print(sp)
  nameSp <- spNames[sp, ]

  spDatasets <- as.data.frame(read_csv(paste(outputFolderSpPreds, "/", sp, "_fullDataset.csv", sep = "")))
  
  spDatasets <- spDatasets %>% tidyterra::select(-1)

  names(spDatasets) <- c(
    "spDensity", "forClass",
    "landClass", "age", "landForClass", "FoLRaster", "sp", "studyArea"
  )
  spDatasets$landForClass <- as.factor(spDatasets$landForClass)
  spDatasets$sp <- as.factor(spDatasets$sp)
  spDatasets$age <- as.integer(spDatasets$age)
  spDatasets$landForClass <- factor(spDatasets$landForClass,
    levels = c(
      "landClass_1",
      "landClass_2",
      "landClass_3",
      "landClass_4",
      "landClass_6",
      "landClass_5",
      "landClass_0",
      "forClass_3",
      "forClass_4",
      "forClass_5",
      "forClass_6",
      "forClass_1",
      "forClass_2"
    )
  )


  densityPlot <- ggplot(spDatasets, aes(x = spDensity)) + 
    geom_density(fill = coloursAge) +
    facet_wrap(~landForClass,
      labeller = labeller(landForClass = class.labs),
      ncol = 4
    ) +
    theme_minimal() +
    # ggtitle(paste("Kernel density plots for ", nameSp, " \nby forest or land cover class", sep = "")) +
    xlab("BAM predicted density (males ha⁻¹)") +
    scale_x_continuous(breaks = pretty_breaks(n = 5), limits = c(0, NA)) +
    ylab("Density") +
    scale_y_continuous(breaks = pretty_breaks(n = 5)) +
    theme(
      # title = element_text(size = 14),
      title = element_blank(),
      strip.text.x = element_text(
        size = 14
      ),
      strip.background = element_rect(
        color = "white", fill = "white", linewidth = 1.1
      ),
      legend.title = element_blank(),
      legend.position = "none",
      axis.text.y = element_text(size = 12),
      axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
      axis.title = element_text(size = 14)
    )

  #save
  ggsave(file.path(saveLocation_1D, paste(sp, "_kernelDensity.jpeg", sep = "")), densityPlot, width = 20, height = 20, units = "cm")
})
```


# PLOTS OF GBM STATS

### Plot relative influence for age and friedman's H statistic

Here a graph of the gbm relative influence for age and cover class is plotted, and the friedman's H statistic for the GBM is plotted per sp.


```{r plotBothRelInfAndFriedmansH, echo=FALSE}
statsGBM <- as.data.table(read_csv(paste(outputFolderSpPreds, "/statsGBM.csv", sep = "")))
statsGBM <- statsGBM[, c("species", "FriedmansHStat", "relInfAge")]
statsGBM <- na.omit(statsGBM)


statsGBM$relInfAge <- statsGBM$relInfAge / 100


plotRelInf <- ggplot(
  data = statsGBM,
  aes(
    x = relInfAge,
    y = species
  )
) +
  geom_point() +
  geom_segment(aes(yend = species, xend = 0)) +
  ggtitle("Relative influence of forest age among species") +
  xlab("Relative influence of age (%)") +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_discrete(labels = c(
    "ALFL" = "Alder Flycatcher",
    "AMGO" = "American Goldfinch",
    "AMRO" = "American Robin",
    "BARS" = "Barn Swallow",
    "BBWA" = "Bay-breasted Warbler",
    "BCCH" = "Black-capped Chickadee",
    "BHCO" = "Brown-headed Cowbird",
    "BOCH" = "Boreal Chickadee",
    "BRBL" = "Brewer's Blackbird",
    "BRCR" = "Brown Creeper",
    "CCSP" = "Clay-colored Sparrow",
    "CEDW" = "Cedar Waxwing",
    "CHSP" = "Chipping Sparrow",
    "CLSW" = "Cliff Swallow",
    "CMWA" = "Cape May Warbler",
    "COYE" = "Common Yellowthroat",
    "DEJU" = "Dark-eyed Junco",
    "GCKI" = "Golden-crowned Kinglet",
    "GRAJ" = "Canada Jay",
    "HETH" = "Hermit Thrush",
    "HOWR" = "House Wren",
    "LEFL" = "Least Flycatcher",
    "MOWA" = "Mourning Warbler",
    "OVEN" = "Ovenbird",
    "PAWA" = "Palm Warbler",
    "RBNU" = "Red-breasted Nuthatch",
    "RCKI" = "Ruby-crowned Kinglet",
    "REVI" = "Red-eyed Vireo",
    "SWTH" = "Swainson’s Thrush",
    "TEWA" = "Tennessee Warbler",
    "YRWA" = "Yellow-rumped Warbler"
  )) +
  theme_minimal() +
  theme(
    title = element_text(size = 16),
    legend.title = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(size = 14),
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_blank(),
    plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm")
  )

plotRelInf


plotFrH <- ggplot(
  data = statsGBM,
  aes(
    x = FriedmansHStat,
    y = species
  )
) +
  geom_point() +
  geom_segment(aes(yend = species, xend = 0)) +
  ggtitle("Friedman's H-statistic by species") +
  xlab("Friedman's H-statistic") +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_discrete(labels = c(
    "ALFL" = "Alder Flycatcher",
    "AMGO" = "American Goldfinch",
    "AMRO" = "American Robin",
    "BARS" = "Barn Swallow",
    "BBWA" = "Bay-breasted Warbler",
    "BCCH" = "Black-capped Chickadee",
    "BHCO" = "Brown-headed Cowbird",
    "BOCH" = "Boreal Chickadee",
    "BRBL" = "Brewer's Blackbird",
    "BRCR" = "Brown Creeper",
    "CCSP" = "Clay-colored Sparrow",
    "CEDW" = "Cedar Waxwing",
    "CHSP" = "Chipping Sparrow",
    "CLSW" = "Cliff Swallow",
    "CMWA" = "Cape May Warbler",
    "COYE" = "Common Yellowthroat",
    "DEJU" = "Dark-eyed Junco",
    "GCKI" = "Golden-crowned Kinglet",
    "GRAJ" = "Canada Jay",
    "HETH" = "Hermit Thrush",
    "HOWR" = "House Wren",
    "LEFL" = "Least Flycatcher",
    "MOWA" = "Mourning Warbler",
    "OVEN" = "Ovenbird",
    "PAWA" = "Palm Warbler",
    "RBNU" = "Red-breasted Nuthatch",
    "RCKI" = "Ruby-crowned Kinglet",
    "REVI" = "Red-eyed Vireo",
    "SWTH" = "Swainson’s Thrush",
    "TEWA" = "Tennessee Warbler",
    "YRWA" = "Yellow-rumped Warbler"
  )) +
  theme_minimal() +
  scale_fill_manual(values = colours2D) +
  theme(
    title = element_text(size = 16),
    legend.title = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(size = 14),
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_blank(),
    plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm")
  )

plotFrH

allMatrixPlot <- ggarrange(
  plotRelInf +
    theme(title = element_blank()),
  plotFrH +
    theme(
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      title = element_blank(),
      legend.position = "none"
    ),
  ncol = 2,
  labels = c("a", "b"),
  label.args = list(gp = gpar(font = 4, fontsize = 16), x = unit(1, "line"), hjust = 1)
)
test <- statsGBM %>%
  rowwise() %>%
  arrange(relInfAge) %>%
  mutate(species = factor(species, species))


plotGBM <- ggplot(test) +
  geom_segment(aes(y = FriedmansHStat, yend = relInfAge, x = species, xend = species), color = "grey") +
  geom_point(aes(x = species, y = relInfAge, colour = "Proportional Influence of Age"), stroke = 1.5, shape = 1, size = 3) +
  geom_point(aes(x = species, y = FriedmansHStat, color = "Friedman's H Statistic"), shape = 16, size = 4) +
  xlab("") +
  ggtitle("Friedman's H-statistic by species \nand the relative influence of Age") +
  scale_color_manual(values = c(colours2D, colours2D_alt)) +
  scale_x_discrete(labels = c(
    "ALFL" = "Alder Flycatcher",
    "AMGO" = "American Goldfinch",
    "AMRO" = "American Robin",
    "BARS" = "Barn Swallow",
    "BBWA" = "Bay-breasted Warbler",
    "BCCH" = "Black-capped Chickadee",
    "BHCO" = "Brown-headed Cowbird",
    "BOCH" = "Boreal Chickadee",
    "BRBL" = "Brewer's Blackbird",
    "BRCR" = "Brown Creeper",
    "CCSP" = "Clay-colored Sparrow",
    "CEDW" = "Cedar Waxwing",
    "CHSP" = "Chipping Sparrow",
    "CLSW" = "Cliff Swallow",
    "CMWA" = "Cape May Warbler",
    "COYE" = "Common Yellowthroat",
    "DEJU" = "Dark-eyed Junco",
    "GCKI" = "Golden-crowned Kinglet",
    "GRAJ" = "Canada Jay",
    "HETH" = "Hermit Thrush",
    "HOWR" = "House Wren",
    "LEFL" = "Least Flycatcher",
    "MOWA" = "Mourning Warbler",
    "OVEN" = "Ovenbird",
    "PAWA" = "Palm Warbler",
    "RBNU" = "Red-breasted Nuthatch",
    "RCKI" = "Ruby-crowned Kinglet",
    "REVI" = "Red-eyed Vireo",
    "SWTH" = "Swainson’s Thrush",
    "TEWA" = "Tennessee Warbler",
    "YRWA" = "Yellow-rumped Warbler"
  )) +
  coord_flip() +
  theme_minimal() +
  theme(
    title = element_blank(),
    legend.title = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(0.2, 0.2, 0.2, 0.2, "cm")
  )

plotGBM

ggsave(file.path(saveLocation, "statsGBM.jpeg"), plotGBM, width = 18, height = 16, units = "cm")
```


# PLOTS AND TABLES OF STATS

## Table of unimodality and normality test results

```{r tabUniAndNormTest, echo=FALSE}
lapply(spList, FUN = function(sp) {
  print(sp)
  nameSp <- spNames[sp, ]
 
  # 1DPS 
  spPreds1D <- as.data.frame(read_csv(paste(outputFolderSpPreds, "/", sp, "_spPreds1D.csv", sep = "")))
  spPreds1D <- spPreds1D[, c("landForClass", "normality_stat", "normality_p", "unimodality_stat", "unimodality_p")]
  print(spPreds1D)
  spPreds1D$landForClass[spPreds1D$landForClass == "forClass_1"] <- "Black Spruce"
  spPreds1D$landForClass[spPreds1D$landForClass == "forClass_2"] <- "Black Spruce Wet"
  spPreds1D$landForClass[spPreds1D$landForClass == "forClass_3"] <- "Deciduous"
  spPreds1D$landForClass[spPreds1D$landForClass == "forClass_4"] <- "Mixed"
  spPreds1D$landForClass[spPreds1D$landForClass == "forClass_5"] <- "Pine"
  spPreds1D$landForClass[spPreds1D$landForClass == "forClass_6"] <- "White Spruce"
  spPreds1D$landForClass[spPreds1D$landForClass == "landClass_0"] <- "Unmanaged Forest"
  spPreds1D$landForClass[spPreds1D$landForClass == "landClass_1"] <- "Water/Ice"
  spPreds1D$landForClass[spPreds1D$landForClass == "landClass_2"] <- "Wetland"
  spPreds1D$landForClass[spPreds1D$landForClass == "landClass_3"] <- "Anthro/Exposed"
  spPreds1D$landForClass[spPreds1D$landForClass == "landClass_4"] <- "Crop/Grass"
  spPreds1D$landForClass[spPreds1D$landForClass == "landClass_5"] <- "Shrub"
  print(spPreds1D)
  spPreds1D$landForClass <- factor(spPreds1D$landForClass, levels = c("Water/Ice", "Wetland", "Anthro/Exposed", "Crop/Grass", "Shrub", "Unmanaged Forest", "Deciduous", "Mixed", "Pine", "White Spruce", "Black Spruce", "Black Spruce Wet"))
  # Define the new order for the Class column
  new_order <- c("Water/Ice", "Wetland", "Anthro/Exposed", "Crop/Grass", "Shrub", "Unmanaged Forest", "Deciduous", "Mixed", "Pine", "White Spruce", "Black Spruce", "Black Spruce Wet")
  # Reorder the rows based on the new order
  spPreds1D <- spPreds1D[match(new_order, spPreds1D$landForClass), ]


  spPreds1D <- COINr::signif_df(spPreds1D, digits = 2)
  print(spPreds1D)
   fileName <- paste(sp, "_NormAndUnimod", sep = "")
  write.csv(spPreds1D, file = file.path(saveLocation, fileName))

  names(spPreds1D) <- c("Class", "D", "p-value ", "Dn", "p-value  ")

  gtTab <- gt::gt(spPreds1D, rowname_col = "Class") |> 
    tab_header(title = paste("Normality and Unimodality Test Results for ", nameSp, sep = "")) |>
    tab_spanner(
      label = "Hartigan's Diptest for Unimodality ",
      columns = c(4:5)
    ) |>
    tab_spanner(
      label = "Anderson-Darling Test for Normality",
      columns = c(2:3)
    ) |>
    tab_row_group(
      label = "Forest Classes",
      rows = c(7:12)
    ) |>
    tab_row_group(
      label = "Land Cover Classes",
      rows = c(1:6)
    )

  gtTab
  gtsave(gtTab, paste(sp, "_NormAndUnimod.png", sep = ""), path = saveLocation_1D, expand = 50)


  
})

```

## Combined unimodality, normality, spearman, morans - different age classes

```{r plotCombinedUnimodSpearAndMoranAgeClass, echo=FALSE}
# get data tables needed

#1DPS
assumptionsByClass1D <- as.data.table(read_csv(paste(outputFolderSpPreds, "/assumptionsByClass1D.csv", sep = "")))
names(assumptionsByClass1D)[names(assumptionsByClass1D) == "landForClass"] <- "landAgeClass"
assumptionsByClass1D <- assumptionsByClass1D[1:6]
assumptionsByClass1D <- assumptionsByClass1D[, c("landAgeClass", "noSps", "propSpsNormal", "propSpsUnimodal", "smoothingType")]
assumptionsByClass1D <- droplevels(assumptionsByClass1D)

#2DPS 10 yr
assumptionsByClass2D_10 <- as.data.frame(read_csv(paste(outputFolderSpPreds_10yr, "/assumptionsByClass2D.csv", sep = "")))
assumptionsByClass2D_10$smoothingType[assumptionsByClass2D_10$smoothingType == "2"] <- "2_10"
assumptionsByClass2D_10 <- assumptionsByClass2D_10[, c("landAgeClass", "noSps", "propSpsNormal", "propSpsUnimodal", "smoothingType")]

#2DPS 20 yr
assumptionsByClass2D_20 <- as.data.frame(read_csv(paste(outputFolderSpPreds_20yr, "/assumptionsByClass2D.csv", sep = "")))
assumptionsByClass2D_20$smoothingType[assumptionsByClass2D_20$smoothingType == "2"] <- "2_20"
assumptionsByClass2D_20 <- assumptionsByClass2D_20[, c("landAgeClass", "noSps", "propSpsNormal", "propSpsUnimodal", "smoothingType")]

#2DPS 40 yr
assumptionsByClass2D_40 <- as.data.frame(read_csv(paste(outputFolderSpPreds_40yr, "/assumptionsByClass2D.csv", sep = "")))
assumptionsByClass2D_40$smoothingType[assumptionsByClass2D_40$smoothingType == "2"] <- "2_40"
assumptionsByClass2D_40 <- assumptionsByClass2D_40[, c("landAgeClass", "noSps", "propSpsNormal", "propSpsUnimodal", "smoothingType")]

assumptionsByClass <- rbind(
  assumptionsByClass1D,
  assumptionsByClass2D_10,
  assumptionsByClass2D_20,
  assumptionsByClass2D_40
)

assumptionsByClass$smoothingType <- as.factor(assumptionsByClass$smoothingType)


# unimodality plot
plotUnimodality <- ggplot(
  data = assumptionsByClass,
  aes(fill = smoothingType, y = propSpsUnimodal, x = smoothingType)
) +
  geom_violin(width = 0.95) +
  annotate("text",
    x = 1:length(table(assumptionsByClass$smoothingType)),
    y = aggregate(propSpsUnimodal ~ smoothingType, assumptionsByClass, max)[, 2],
    label = paste("n = ", table(assumptionsByClass$smoothingType), sep = ""),
    size = 3,
    vjust = -1
  ) +
  ggtitle("Proportion of bird species per class \nwith a unimodal distribution (p < 0.05)") +
  xlab("Species") +
  ylab("Proportion of bird \nspecies unimodal") +
  scale_y_continuous(breaks = pretty_breaks(), limits = c(NA, 1)) +
  theme_classic() +
  scale_x_discrete(labels = c(
    "1" = "1DPS",
    "2_10" = "2DPS (10-year\n age classes)",
    "2_20" = "2DPS (20-year\n age classes)",
    "2_40" = "2DPS (40-year\n age classes)"
  )) +
  scale_fill_manual(values = coloursAllSame) +
  theme(
    title = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "none",
    legend.text = element_blank(),
    axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 9),
    axis.title.y = element_text(size = 10),
    axis.title.x = element_blank(),
    plot.margin = margin(0.5, 0.5, 0, 0.5, "cm"),
    axis.line.y = element_line(linewidth = 1)
  )

plotUnimodality


# normality plot
plotNormality <- ggplot(
  data = assumptionsByClass,
  aes(fill = smoothingType, y = propSpsNormal, x = smoothingType)
) +
  geom_violin(width = 0.95) +
  annotate("text",
    x = 1:length(table(assumptionsByClass$smoothingType)),
    y = aggregate(propSpsNormal ~ smoothingType, assumptionsByClass, max)[, 2],
    label = paste("n = ", table(assumptionsByClass$smoothingType), sep = ""),
    size = 3,
    vjust = -1
  ) +
  ggtitle("Proportion of bird species per class \nwith a normal distribution (p < 0.05)") +
  xlab("Species") +
  ylab("Proportion of bird \nspecies normal") +
  scale_y_continuous(breaks = pretty_breaks(), limits = c(NA, 0.35)) +
  theme_classic() +
  scale_x_discrete(labels = c(
    "1" = "1DPS",
    "2_10" = "2DPS (10-year\n age classes)",
    "2_20" = "2DPS (20-year\n age classes)",
    "2_40" = "2DPS (40-year\n age classes)"
  )) +
  scale_fill_manual(values = coloursAllSame) +
  theme(
    title = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "none",
    legend.text = element_blank(),
    axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 9),
    axis.title.y = element_text(size = 10),
    axis.title.x = element_blank(),
    plot.margin = margin(0.5, 0.5, 0, 0.5, "cm"),
    axis.line.y = element_line(linewidth = 1)
  )

plotNormality

##################################################################

#    spearmanStats
spearmanStats_10 <- as.data.frame(read_csv(paste(outputFolderSpPreds_10yr, "/spearmanStats.csv", sep = "")))
names(spearmanStats_10) <- c("species", "spearman1D", "spearman2D_10")

spearmanStats_20 <- as.data.frame(read_csv(paste(outputFolderSpPreds_20yr, "/spearmanStats.csv", sep = "")))
names(spearmanStats_20) <- c("species", "spearman1D", "spearman2D_20")

spearmanStats_40 <- as.data.frame(read_csv(paste(outputFolderSpPreds_40yr, "/spearmanStats.csv", sep = "")))
names(spearmanStats_40) <- c("species", "spearman1D", "spearman2D_40")

spearmanStats <- cbind(
  spearmanStats_10,
  spearmanStats_20$spearman2D_20,
  spearmanStats_40$spearman2D_40
)
names(spearmanStats) <- c("species", "spearman1D", "spearman2D_10", "spearman2D_20", "spearman2D_40")
spearmanStats <- spearmanStats %>%
  pivot_longer(
    cols = c(spearman1D, spearman2D_10, spearman2D_20, spearman2D_40), 
    names_to = "smoothingType", 
    values_to = "spearmanStat" )


plotSpearmanStats <- ggplot(
  data = spearmanStats,
  aes(
    y = spearmanStat,
    x = smoothingType,
    fill = smoothingType
  )
) +
  geom_violin(width = 0.95) +
  geom_boxplot(width = 0.15, color = "grey") +
  ggtitle("Correlation between BAM national models \nand mapped predictions, across sp species") +
  scale_x_discrete(labels = c(
    "spearman1D" = "1DPS",
    "spearman2D_10" = "2DPS (10-year\n age classes)",
    "spearman2D_20" = "2DPS (20-year\n age classes)",
    "spearman2D_40" = "2DPS (40-year\n age classes)"
  )) +
  ylab("Spearman's \ncoefficient (p)") +
  scale_y_continuous(breaks = pretty_breaks()) +
  theme_classic() +
  scale_fill_manual(
    values = coloursAllSame,
    labels = c("1D Piecewise\n Smoothing", "2D Piecewise\n Smoothing")
  ) +
  theme(
    title = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "none",
    legend.text = element_text(size = 10),
    axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 9),
    axis.title.y = element_text(size = 10),
    axis.title.x = element_blank(),
    plot.margin = margin(0.2, 0.2, 0.2, 0.2, "cm")
  )

plotSpearmanStats


#################################################################

# resStats
resStats_10 <- as.data.frame(read_csv(paste(outputFolderSpPreds_10yr, "/resStats.csv", sep = "")))
names(resStats_10) <- c("species", "medianRes1D", "medianRes2D_10", "moran1D", "moran2D_10")

resStats_20 <- as.data.frame(read_csv(paste(outputFolderSpPreds_20yr, "/resStats.csv", sep = "")))
names(resStats_20) <- c("species", "medianRes1D", "medianRes2D_20", "moran1D", "moran2D_20")

resStats_40 <- as.data.frame(read_csv(paste(outputFolderSpPreds_40yr, "/resStats.csv", sep = "")))
names(resStats_40) <- c("species", "medianRes1D", "medianRes2D_40", "moran1D", "moran2D_40")

resStats <- as.data.frame(cbind(
  resStats_10$species,
  resStats_10$moran1D,
  resStats_10$moran2D_10,
  resStats_20$moran2D_20,
  resStats_40$moran2D_40
))
names(resStats) <- c("species", "moran1D", "moran2D_10", "moran2D_20", "moran2D_40")
resStats$species <- as.factor(resStats$species)
resStats$moran1D <- as.numeric(resStats$moran1D)
resStats$moran2D_10 <- as.numeric(resStats$moran2D_10)
resStats$moran2D_20 <- as.numeric(resStats$moran2D_20)
resStats$moran2D_40 <- as.numeric(resStats$moran2D_40)


## moran stats plot

moranStats <- resStats %>%
  pivot_longer(
    cols = c(moran1D, moran2D_10, moran2D_20, moran2D_40), 
    names_to = "smoothingType", 
    values_to = "MoransI" 
  )

plotAutocorStats <- ggplot(
  data = moranStats,
  aes(
    fill = factor(smoothingType,
      levels = c(
        "moran1D",
        "moran2D_10",
        "moran2D_20",
        "moran2D_40"
      )
    ),
    y = MoransI,
    x = smoothingType
  )
) +
  geom_violin(width = 0.95) +
  geom_boxplot(width = 0.15, color = "grey") +
  ggtitle("Spatial autocorrelation of mapped residuals, \nacross bird species") +
  ylab("Moran's I") +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_discrete(labels = c(
    "moran1D" = "1DPS",
    "moran2D_10" = "2DPS (10-year\n age classes)",
    "moran2D_20" = "2DPS (20-year\n age classes)",
    "moran2D_40" = "2DPS (40-year\n age classes)"
  )) +
  theme_classic() +
  scale_fill_manual(
    values = coloursAllSame,
    labels = c("moran1D", "moran2D")
  ) +
  theme(
    title = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "none",
    legend.text = element_text(size = 10),
    axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 9),
    axis.title.y = element_text(size = 10),
    axis.title.x = element_blank(),
    plot.margin = margin(0.2, 0.2, 0.2, 0.2, "cm")
  )

plotAutocorStats

#################################################################


allMatrixPlot <- ggarrange(
  plotUnimodality +
    theme(
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      title = element_blank(),
      legend.position = "none"
    ),
  plotNormality +
    theme(
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      title = element_blank(),
      legend.position = "none"
    ),
  plotSpearmanStats +
    theme(
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      title = element_blank(),
      legend.position = "none"
    ),
  plotAutocorStats +
    theme(title = element_blank()),
  ncol = 1,
  labels = c("(a)", "(b)", "(c)", "(d)"),
  label.args = list(gp = gpar(font = 4, fontsize = 12), x = unit(1, "line"), hjust = 1)
)

ggsave(file.path(saveLocation_1D, "unimodSpearmanAutocor.jpeg"), allMatrixPlot, width = 14, height = 25, units = "cm")


############################################################################


plotUniNorm <- ggarrange(
  plotUnimodality +
    theme(
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      title = element_blank(),
      legend.position = "none"
    ),
  plotNormality +
    theme(title = element_blank()),
  ncol = 1,
  labels = c("(a)", "(b)"),
  label.args = list(gp = gpar(font = 4, fontsize = 12), x = unit(1, "line"), hjust = 1)
)

ggsave(file.path(saveLocation_1D, "plotUniNorm.jpeg"), plotUniNorm, width = 10, height = 12, units = "cm")


#########################################################################################


plotSpearAuto <- ggarrange(
  plotSpearmanStats +
    theme(
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      title = element_blank(),
      legend.position = "none"
    ),
  plotAutocorStats +
    theme(title = element_blank()),
  ncol = 1,
  labels = c("(a)", "(b)"),
  label.args = list(gp = gpar(font = 4, fontsize = 12), x = unit(1, "line"), hjust = 1)
)

ggsave(file.path(saveLocation_1D, "plotSpearAuto.jpeg"), plotSpearAuto, width = 10, height = 12, units = "cm")
```



##Tables of test stats

```{r tabCombinedUnimodSpearAndMoranAgeClass, echo=FALSE}
#    spearmanStats
spearmanStats_10 <- as.data.frame(read_csv(paste(outputFolderSpPreds_10yr, "/spearmanStats.csv", sep = "")))
names(spearmanStats_10) <- c("species", "spearman1D", "spearman2D_10")

spearmanStats_20 <- as.data.frame(read_csv(paste(outputFolderSpPreds_20yr, "/spearmanStats.csv", sep = "")))
names(spearmanStats_20) <- c("species", "spearman1D", "spearman2D_20")

spearmanStats_40 <- as.data.frame(read_csv(paste(outputFolderSpPreds_40yr, "/spearmanStats.csv", sep = "")))
names(spearmanStats_40) <- c("species", "spearman1D", "spearman2D_40")

spearmanStats <- cbind(
  spNames$spName, spearmanStats_10,
  spearmanStats_20$spearman2D_20,
  spearmanStats_40$spearman2D_40
)

spearmanStats$species <- NULL

names(spearmanStats) <- c("Bird Species", "1DPS  ", "2DPS (10-year age classes)", "2DPS (20-year age classes)", "2DPS (40-year age classes)")

tabSpearman <- gt::gt(spearmanStats) |> 
  tab_header(title = "Spearman's Rank Correlation Test Statistic (ρ)") |>
  tab_stubhead(label = "Bird Species")

gtsave(tabSpearman, "spearmanStatsTab.png", path = saveLocation_1D, expand = 50)


# resStats
resStats_10 <- as.data.frame(read_csv(paste(outputFolderSpPreds_10yr, "/resStats.csv", sep = "")))
names(resStats_10) <- c("species", "medianRes1D", "medianRes2D_10", "moran1D", "moran2D_10")

resStats_20 <- as.data.frame(read_csv(paste(outputFolderSpPreds_20yr, "/resStats.csv", sep = "")))
names(resStats_20) <- c("species", "medianRes1D", "medianRes2D_20", "moran1D", "moran2D_20")

resStats_40 <- as.data.frame(read_csv(paste(outputFolderSpPreds_40yr, "/resStats.csv", sep = "")))
names(resStats_40) <- c("species", "medianRes1D", "medianRes2D_40", "moran1D", "moran2D_40")

moranStats <- as.data.frame(cbind(
  spNames$spName,
  resStats_10$species,
  resStats_10$moran1D,
  resStats_10$moran2D_10,
  resStats_20$moran2D_20,
  resStats_40$moran2D_40
))
names(moranStats) <- c("species", "spCode", "moran1D", "moran2D_10", "moran2D_20", "moran2D_40")
moranStats$species <- as.factor(moranStats$species)
moranStats$spCode <- as.factor(moranStats$spCode)
moranStats$moran1D <- as.numeric(moranStats$moran1D)
moranStats$moran2D_10 <- as.numeric(moranStats$moran2D_10)
moranStats$moran2D_20 <- as.numeric(moranStats$moran2D_20)
moranStats$moran2D_40 <- as.numeric(moranStats$moran2D_40)

moranStats$spCode <- NULL
names(moranStats) <- c("Bird Species", "1DPS  ", "2DPS (10-year age classes)", "2DPS (20-year age classes)", "2DPS (40-year age classes)")

## moran stats plot
tabMoran <- gt::gt(moranStats) |> 
  tab_header(title = "Moran's I") |>
  tab_stubhead(label = "Bird Species")

gtsave(tabMoran, "moranStatsTab.png", path = saveLocation_1D, expand = 50)
```



# RASTER VISUALISATION 

## Set of three rasters for one species

```{r plotExampleBirdRasters, echo=FALSE}


lapply(spList, FUN = function(sp) {
  
  print(sp)

  nameSp <- spNames[sp, ]
  nmRas <- terra::rast(paste(locationSpRas, "/", sp, "-meanBoot_BCR-60_studyArea_AB_BCR6", sep = ""))
  mapRas <- terra::rast(paste(outputFolderSpPredsRasters, "/", sp, "_for2DAndLc1DMap", sep = ""))
  resRas <- terra::rast(paste(outputFolderSpPredsRasters, "/", sp, "_for2DAndLc1DRes", sep = ""))

  allRasters <- c(nmRas, mapRas, resRas)
  names(allRasters) <- c("Species Density Raster", "2DPS Prediction", "Residual")

  studyArea <- terra::project(studyArea, y = "+proj=longlat +datum=WGS84")
  allRasters <- terra::project(allRasters, y = "+proj=longlat +datum=WGS84")
  allRasters <- crop(allRasters, studyArea)

  rasterPlot <- ggplot() +
    geom_spatraster(data = allRasters) +
    facet_wrap(~lyr, ncol = 3) +
    scale_fill_gradient2(
      low = "violetred4",
      mid = "white",
      high = "darkolivegreen",
      midpoint = 0,
      na.value = "white"
    ) +
    theme_classic() +
    labs(title = paste(" ", nameSp, " density predictions (males per ha)", sep = "")) +
    theme(
      title = element_text(size = 14),
      strip.text.x = element_text(
        size = 12
      ),
      strip.text.y = element_text(
        size = 11
      ),
      legend.position = "right",
      legend.text = element_text(size = 10),
      legend.title = element_blank(),
      legend.key.height = unit(1, "null"),
      strip.background = element_rect(
        color = "white", fill = "white", linewidth = 0.1
      ),
      axis.line = element_blank(),
      axis.ticks = element_blank(),
      axis.text.y = element_blank(),
      axis.text.x = element_blank(),
      axis.title.y = element_blank(),
      axis.title.x = element_text(size = 11)
    )

  ggsave(file.path(paste(saveLocation, "/", sp, "_rasterPlots.jpeg", sep = "")), rasterPlot, width = 24, height = 14, units = "cm")
})
```



##All rasters per species - all age classes

```{r plotExampleSpRasters, echo=FALSE}

lapply(spList, FUN = function(sp) {
  
  print(sp)

  nameSp <- spNames[sp, ]
  nmRas <- terra::rast(paste(locationSpRas, "/", sp, "-meanBoot_BCR-60_studyArea_AB_BCR6", sep = ""))
  mapRas_1D <- terra::rast(paste(outputFolderSpPredsRasters_10yr, "/", sp, "_for1DAndLc1DMap", sep = ""))
  resRas_1D <- terra::rast(paste(outputFolderSpPredsRasters_10yr, "/", sp, "_for1DAndLc1DRes", sep = ""))

  mapRas_2D_10 <- terra::rast(paste(outputFolderSpPredsRasters_10yr, "/", sp, "_for2DAndLc1DMap", sep = ""))
  resRas_2D_10 <- terra::rast(paste(outputFolderSpPredsRasters_10yr, "/", sp, "_for2DAndLc1DRes", sep = ""))

  mapRas_2D_20 <- terra::rast(paste(outputFolderSpPredsRasters_20yr, "/", sp, "_for2DAndLc1DMap", sep = ""))
  resRas_2D_20 <- terra::rast(paste(outputFolderSpPredsRasters_20yr, "/", sp, "_for2DAndLc1DRes", sep = ""))

  mapRas_2D_40 <- terra::rast(paste(outputFolderSpPredsRasters_40yr, "/", sp, "_for2DAndLc1DMap", sep = ""))
  resRas_2D_40 <- terra::rast(paste(outputFolderSpPredsRasters_40yr, "/", sp, "_for2DAndLc1DRes", sep = ""))

  allRasters <- c(nmRas, mapRas_1D, resRas_1D, 
                  mapRas_2D_10, resRas_2D_10,
                  mapRas_2D_20, resRas_2D_20,  
                  mapRas_2D_40, resRas_2D_40)
  names(allRasters) <- c(
    "BAM Model Raster", "1DPS Prediction", "1DPS Residual",
    "2DPS Prediction (10-year age classes)", "2DPS Residual (10-year age classes)",
     "2DPS Prediction (20-year age classes)", "2DPS Residual (20-year age classes)",
     "2DPS Prediction (40-year age classes)", "2DPS Residual (40-year age classes)"
  )

  studyArea <- terra::project(studyArea, y = "+proj=longlat +datum=WGS84")
  allRasters <- terra::project(allRasters, y = "+proj=longlat +datum=WGS84")
  allRasters <- crop(allRasters, studyArea)
  allRasAsDF <- as.data.frame(allRasters, xy = TRUE, cells = TRUE)

  allRasAsDF <- pivot_longer(allRasAsDF,
    cols = c(4:12),
    names_to = "RasterName", 
    values_to = "value"
  )
  allRasAsDF <- allRasAsDF %>%
    mutate(rasterType = case_when(
      RasterName == "BAM Model Raster" ~ "BAM Model",
      RasterName == "1DPS Prediction" ~ "Piecewise Smoothing",
        RasterName == "2DPS Prediction (10-year age classes)" ~ "Piecewise Smoothing",
      RasterName == "2DPS Prediction (20-year age classes)" ~ "Piecewise Smoothing",
      RasterName == "2DPS Prediction (40-year age classes)" ~ "Piecewise Smoothing",
      RasterName == "1DPS Residual" ~ "Residual",
        RasterName == "2DPS Residual (10-year age classes)" ~ "Residual",
       RasterName == "2DPS Residual (20-year age classes)" ~ "Residual",
       RasterName == "2DPS Residual (40-year age classes)" ~ "Residual",
    )) %>%
    mutate(smoothingType = case_when(
      RasterName == "BAM Model Raster" ~ "1DPS",
      RasterName == "1DPS Prediction" ~ "1DPS",
      RasterName == "2DPS Prediction (10-year age classes)" ~ "2DPS \n(10-year age classes)",
      RasterName == "2DPS Prediction (20-year age classes)" ~ "2DPS \n(20-year age classes)",
      RasterName == "2DPS Prediction (40-year age classes)" ~ "2DPS \n(40-year age classes)",
      RasterName == "1DPS Residual" ~ "1DPS",
      RasterName == "2DPS Residual (10-year age classes)" ~ "2DPS \n(10-year age classes)",
      RasterName == "2DPS Residual (20-year age classes)" ~ "2DPS \n(20-year age classes)",
      RasterName == "2DPS Residual (40-year age classes)" ~ "2DPS \n(40-year age classes)",
    ))
 
  #make plot
  rasterPlot <- ggplot(allRasAsDF) +
    geom_raster(aes(x = x, y = y, fill = value)) +
    facet_grid(smoothingType ~ rasterType) +
    scale_fill_gradient2(
      mid = "white",
      midpoint = 0,
      low = "violetred4",
      high = "darkolivegreen",
      na.value = "white"
    ) +
    theme_classic() +
    labs(title = paste("Predicted density (males ha⁻¹)", sep = "")) +
    theme(
      title = element_text(size = 16),
      strip.text.x = element_text(
        size = 16
      ),
      strip.text.y = element_text(
        size = 16
      ),
      legend.position = "right",
      legend.text = element_text(size = 16),
      legend.title = element_blank(),
      legend.key.height = unit(1, "null"),
      strip.background = element_rect(
        color = "white", fill = "white", linewidth = 0.1
      ),
      axis.line = element_blank(),
      axis.ticks = element_blank(),
      axis.text.y = element_blank(),
      axis.text.x = element_blank(),
      axis.title.y = element_blank(),
      axis.title.x = element_blank()
    )

  rasterPlot

  #save
  ggsave(file.path(paste(saveLocation_1D, "/", sp, "_rasterPlots.jpeg", sep = "")), rasterPlot, width = 22, height = 35, units = "cm")
  
})
```
